### 2020-07-21 code/figure2.Rmd
updated code to work with new cluster assignments by simon

<div style="text-align: right"> Figure 2 Data </div>
```{r oncoprint-top100-plus-panel-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_2d_3d_coors.tsv") ## load in data
load("../data/gdc_reactome_path_analysis.Rda")

@@

for(n in 1:12) {
  class_spec_sample_id <- tumor_clusters %>% .[,c("V1", "knn_clust")] %>% filter(knn_clust == n) %>% .[,.$V1] ## get class spec sample ids
  top10path <- gdc_reactome_path_analysis %>% .[, names(.) %in% class_spec_sample_id] %>% rowSums() %>% as.data.frame() %>% rownames_to_column(var = "path_name") %>% arrange(desc(.)) %>% .[1:10,] %>% .[, "path_name"] ## getting the top10 mutated pathways in class
  top10path <- gdc_reactome_path_analysis[top10path, class_spec_sample_id]
  top10path[top10path == "1"] <- "MUT" ## categorize boolean values
  top10path[top10path == "0"] <- " "
  col <- c("MUT" = "firebrick3")
heatmap_legend <- list(title = "", at = c("MUT"),
                       labels = c("Mutation"))
alter_fun = list(
    background = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), 
            gp = gpar(fill = "#CCCCCC", col = NA))
    },
    # bug red
    MUT = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), 
            gp = gpar(fill = col["MUT"], col = NA))
    })
  column_title <- "test"
  onco_res <- oncoPrint(top10path,
    alter_fun = alter_fun, col = col,
    column_title = column_title, heatmap_legend_param = heatmap_legend,
    remove_empty_columns = F, remove_empty_rows = F, right_annotation = NULL,
    top_annotation = NULL
    )
  clust_col_order[[n]] <- names(top10path[,onco_res@column_order])
  path_row_order[[n]] <- rownames(top10path[onco_res@row_order,])
}

@@

plots_onco <- list() ## creating a list to store all the ggplots in the for loop below
onco_color <- c("0" = "grey", "1" = "firebrick3") ## color scheme for oncoprint
figure_labs <- LETTERS[seq(from = 2, to = 13)]
for(n in 1:12) {
  onco_ggplot <- gdc_reactome_path_analysis %>% .[rownames(.) %in% path_row_order[[n]],] %>% ## only include the top 10 paths
  .[,names(.) %in% clust_col_order[[n]]] %>% ## only include specific tumor class
  rownames_to_column(var = "path_name") %>% ## path_name to col
  gather(., key = "sample_id", value = "mutation_status", -path_name) %>% ## wide to long format
  mutate(path_name = factor(.$path_name, levels = c(rev(path_row_order[[n]]))), ## factor to force oncoprint order
         sample_id = factor(.$sample_id, levels = c(clust_col_order[[n]])))
  plots_onco[[n]] <- ggplot(data = onco_ggplot, aes(x=sample_id, y=path_name)) + ## ggplot code
  geom_tile(aes(fill = as.character(mutation_status))) +
  scale_fill_manual(values = onco_color) +
  xlab(label = paste0("Class ", n)) +
  labs(tag = figure_labs[n]) +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.position = "none") 
}

@@

fig2A + plots_onco[[1]] + plots_onco[[2]] + plots_onco[[3]] + plots_onco[[4]] + plots_onco[[5]] + plots_onco[[6]] + plots_onco[[7]] + plots_onco[[8]] + plots_onco[[9]] + plots_onco[[10]] + plots_onco[[11]] + plots_onco[[12]] + plot_layout(design = layout)
```

>>>>

<div style="text-align: right"> Figure 2 Data </div>
```{r oncoprint-top100-plus-panel-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_3d_coors.tsv") ## load in data
load("../data/gdc_reactome_path_analysis.Rda")

@@

for(n in 1:max(tumor_clusters$clust_knn)) {
  class_spec_sample_id <- tumor_clusters %>% .[,c("sample_id", "clust_knn")] %>% filter(clust_knn == n) %>% .[,.$sample_id] ## get class spec sample ids
  top10path <- gdc_reactome_path_analysis %>% .[, names(.) %in% class_spec_sample_id] %>% rowSums() %>% as.data.frame() %>% rownames_to_column(var = "path_name") %>% arrange(desc(.)) %>% .[1:10,] %>% .[, "path_name"] ## getting the top10 mutated pathways in class
  top10path <- gdc_reactome_path_analysis[top10path, class_spec_sample_id]
  top10path[top10path == "1"] <- "MUT" ## categorize boolean values
  top10path[top10path == "0"] <- " "
  col <- c("MUT" = "firebrick3")
heatmap_legend <- list(title = "", at = c("MUT"),
                       labels = c("Mutation"))
alter_fun = list(
    background = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), 
            gp = gpar(fill = "#CCCCCC", col = NA))
    },
    # bug red
    MUT = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), 
            gp = gpar(fill = col["MUT"], col = NA))
    })
  column_title <- "test"
  onco_res <- oncoPrint(top10path,
    alter_fun = alter_fun, col = col,
    column_title = column_title, heatmap_legend_param = heatmap_legend,
    remove_empty_columns = F, remove_empty_rows = F, right_annotation = NULL,
    top_annotation = NULL
    )
  clust_col_order[[n]] <- names(top10path[,onco_res@column_order])
  path_row_order[[n]] <- rownames(top10path[onco_res@row_order,])
}

@@

plots_onco <- list() ## creating a list to store all the ggplots in the for loop below
onco_color <- c("0" = "grey", "1" = "firebrick3") ## color scheme for oncoprint
for(n in 1:length(clust_col_order)) {
  onco_ggplot <- gdc_reactome_path_analysis %>% .[rownames(.) %in% path_row_order[[n]],] %>% ## only include the top 10 paths
  .[,names(.) %in% clust_col_order[[n]]] %>% ## only include specific tumor class
  rownames_to_column(var = "path_name") %>% ## path_name to col
  gather(., key = "sample_id", value = "mutation_status", -path_name) %>% ## wide to long format
  mutate(path_name = factor(.$path_name, levels = c(rev(path_row_order[[n]]))), ## factor to force oncoprint order
         sample_id = factor(.$sample_id, levels = c(clust_col_order[[n]])))
  plots_onco[[n]] <- ggplot(data = onco_ggplot, aes(x=sample_id, y=path_name)) + ## ggplot code
  geom_tile(aes(fill = as.character(mutation_status))) +
  scale_fill_manual(values = onco_color) +
  xlab(label = paste0("Class ", n)) +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.position = "none") 
}

@@

fig2A + plots_onco[[1]] + plots_onco[[2]] + plots_onco[[3]] + plots_onco[[4]] + plots_onco[[5]] + plots_onco[[6]] + plots_onco[[7]] + plots_onco[[8]] + plots_onco[[9]] + plots_onco[[10]] + plot_layout(design = layout) + plot_annotation(tag_levels = "A")
```

### 2020-07-21 code/figure4.Rmd
updated code to work with new cluster assignments by simon

```{r figure-4-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_2d_3d_coors.tsv") ## load in data
load("../data/gene2_level2_lookup.Rda")
load("../data/pathname2level2.Rda")
load("../data/gdc_reactome_path_analysis.Rda")
load("../data/moderate_enrich_prob.Rda")
load("../data/gene_data.Rda")

## Parameters ##
filtered_class <- c(6,7) ## filtering for only class 6 and 7 samples
n_path <- 10 ## looking at the top10 path
enrich_cutoff <- 0.30
################

>>>>

```{r figure-4-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_3d_coors.tsv") ## load in data
load("../data/gene2_level2_lookup.Rda")
load("../data/pathname2level2.Rda")
load("../data/gdc_reactome_path_analysis.Rda")
load("../data/moderate_enrich_prob.Rda")
load("../data/gene_data.Rda")

## Parameters ##
filtered_class <- c(2) ## filtering for only class 2 samples
n_path <- 10 ## looking at the top10 path
enrich_cutoff <- 0.30
################

### 2020-07-21 code/table1.Rmd
updated code to work with new cluster assignments by simon

<div style="text-align: right"> Table 1 Data </div>
```{r table-1-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_2d_3d_coors.tsv") ## load in data
load("../data/moderate_enrich_prob.Rda") ## load data
load("../data/gene2_level2_lookup.Rda")
load("../data/gene_data.Rda")
load("../data/gdc_gene_exp_adj.Rda")
load("../data/pathname2level2.Rda")

@@

for(d in unique(clust_spec_enrich_top5$clust)) {
  message(d)
  clust_toppath <- clust_spec_enrich_top5 %>% filter(clust == d) %>% .$path %>% as.character() ## get the top pathway names for spec class
  class_spec <- tumor_clusters %>% dplyr::select(V1, knn_clust, project_code) %>% filter(knn_clust %in% d) ## get the smaple ids and cancer type for spec class
  top_genes_in_path_cancer_type <- list()
  for(n in clust_toppath) { ## for each of the top paths do this
    message(n)
    gene_in_path <- id_to_names %>% .[,which(names(.) == n)] %>% as.data.frame() %>% ## select pathway
    'rownames<-' (rownames(id_to_names)) %>% 
    filter(. == 1) %>% ## filter for genes that are in that pathway
    rownames(.)
    top_gene_in_path <- list()
    for(t in type) { ## looping through cancer type within class
      message(t)
      class_spec <- tumor_clusters %>% dplyr::select(V1, knn_clust, project_code) %>% filter(knn_clust == d) ## sample_id for class
      if(t %in% class_spec$project_code) { ## if cancer type is in class then run this 
        class_type_spec <- class_spec %>% filter(project_code == t) %>% column_to_rownames(var = "V1") %>% rownames() ## get cancer type specific tumor classes
      if(length(class_type_spec) == 1) { ## if you only have 1 sample_id in the class
        top_mut_gene <- gene_mutation %>% rownames_to_column(var = "gene") %>% .[.$gene %in% gene_in_path, names(.) %in% c("gene",  class_type_spec)] %>% remove_rownames() %>% column_to_rownames(var = "gene") %>% 'colnames<-' ("obs_count") %>% 
          mutate(n_sample = length(class_type_spec),
               perc_mut = round((obs_count / length(class_type_spec)) * 100, digits = 1)) %>% 
        rownames_to_column(var = "gene") %>% arrange(desc(obs_count))
      }
      if(length(class_type_spec) > 1) { ## if you have more than 1 sample_id in the class
        top_mut_gene <- gene_mutation %>% .[rownames(.) %in% gene_in_path, names(.) %in% class_type_spec] %>% 
        rowSums() %>% as.data.frame() %>% 'colnames<-' ("obs_count") %>%
        mutate(n_sample = length(class_type_spec),
               perc_mut = round((obs_count / length(class_type_spec)) * 100, digits = 1)) %>% 
        rownames_to_column(var = "gene") %>% arrange(desc(obs_count))
      }
      if(show_perc == TRUE & show_n == FALSE) {
         top_gene_in_path[[t]] <- paste0(top_mut_gene$gene[mutated_gene_position], " (", top_mut_gene$perc_mut[mutated_gene_position],"%)")  
      }
      if(show_perc == TRUE & show_n == TRUE) {
         top_gene_in_path[[t]] <- paste0(top_mut_gene$gene[mutated_gene_position], " (", top_mut_gene$perc_mut[mutated_gene_position],"%)", "    n = ", length(class_type_spec))  
      }
      if(show_perc == FALSE) {
         top_gene_in_path[[t]] <- paste0(top_mut_gene$gene[mutated_gene_position])  
      }
      }
      if(!(t %in% class_spec$project_code)){ ## if cancer type is not in class then insert NA
        top_gene_in_path[[t]] <- "NA" 
      }
    }
    top_gene_in_path <- top_gene_in_path %>% do.call(rbind, .) %>% as.data.frame() %>% 'colnames<-' (n)
    clust_row <- data.frame(Class = d, row.names = n) %>% t() ## creating a row for cluster membership
    top_genes_in_path_cancer_type[[n]] <- rbind(clust_row, top_gene_in_path)
  }
  cluster_spec_enrich_table[[n]] <- top_genes_in_path_cancer_type %>% do.call(cbind, .) %>% as.data.frame() %>% 
  t() %>% as.data.frame() %>% rownames_to_column(var = "Pathway")
  }

cluster_spec_enrich_table1 <- cluster_spec_enrich_table %>% do.call(rbind, .) %>% as.data.frame()
save(cluster_spec_enrich_table1, file = "cluster_spec_enrich_table1.Rda")
rm(list = ls(all.names = T))
gc()
```

```{r table-1}
load("../data/cluster_spec_enrich_table1.Rda")
cluster_spec_enrich_table1 %>% kable(caption = "Table 1: Most mutated gene in class-specific pathways (enrichment score >= 0.30) for specific cancer", row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, font_size = 12) %>%
  column_spec(1, bold = T)
```

>>>>

<div style="text-align: right"> Table 1 Data </div>
```{r table-1-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_3d_coors.tsv") ## load in data
load("../data/moderate_enrich_prob.Rda") ## load data
load("../data/gene2_level2_lookup.Rda")
load("../data/gene_data.Rda")
load("../data/gdc_gene_exp_adj.Rda")
load("../data/pathname2level2.Rda")

@@

for(d in unique(clust_spec_enrich_top5$clust)) {
  message(d)
  clust_toppath <- clust_spec_enrich_top5 %>% filter(clust == d) %>% .$path %>% as.character() ## get the top pathway names for spec class
  class_spec <- tumor_clusters %>% dplyr::select(sample_id, clust_knn, project_code) %>% filter(clust_knn %in% d) ## get the smaple ids and cancer type for spec class
  top_genes_in_path_cancer_type <- list()
  for(n in clust_toppath) { ## for each of the top paths do this
    message(n)
    gene_in_path <- id_to_names %>% .[,which(names(.) == n)] %>% as.data.frame() %>% ## select pathway
    'rownames<-' (rownames(id_to_names)) %>% 
    filter(. == 1) %>% ## filter for genes that are in that pathway
    rownames(.)
    top_gene_in_path <- list()
    for(t in type) { ## looping through cancer type within class
      message(t)
      class_spec <- tumor_clusters %>% dplyr::select(sample_id, clust_knn, project_code) %>% filter(clust_knn == d) ## sample_id for class
      if(t %in% class_spec$project_code) { ## if cancer type is in class then run this 
        class_type_spec <- class_spec %>% filter(project_code == t) %>% column_to_rownames(var = "sample_id") %>% rownames() ## get cancer type specific tumor classes
      if(length(class_type_spec) == 1) { ## if you only have 1 sample_id in the class
        top_mut_gene <- gene_mutation %>% rownames_to_column(var = "gene") %>% .[.$gene %in% gene_in_path, names(.) %in% c("gene",  class_type_spec)] %>% remove_rownames() %>% column_to_rownames(var = "gene") %>% 'colnames<-' ("obs_count") %>%
          rownames_to_column(var = "gene") %>%
          mutate(n_sample = length(class_type_spec),
               perc_mut = round((obs_count / length(class_type_spec)) * 100, digits = 1)) %>% 
         arrange(desc(obs_count))
      }
      if(length(class_type_spec) > 1) { ## if you have more than 1 sample_id in the class
        top_mut_gene <- gene_mutation %>% .[rownames(.) %in% gene_in_path, names(.) %in% class_type_spec] %>% 
        rowSums() %>% as.data.frame() %>% 'colnames<-' ("obs_count") %>%
        rownames_to_column(var = "gene") %>%
        mutate(n_sample = length(class_type_spec),
               perc_mut = round((obs_count / length(class_type_spec)) * 100, digits = 1)) %>% 
         arrange(desc(obs_count))
      }
      if(show_perc == TRUE & show_n == FALSE) {
         top_gene_in_path[[t]] <- paste0(top_mut_gene$gene[mutated_gene_position], " (", top_mut_gene$perc_mut[mutated_gene_position],"%)")  
      }
      if(show_perc == TRUE & show_n == TRUE) {
         top_gene_in_path[[t]] <- paste0(top_mut_gene$gene[mutated_gene_position], " (", top_mut_gene$perc_mut[mutated_gene_position],"%)", "    n = ", length(class_type_spec))  
      }
      if(show_perc == FALSE) {
         top_gene_in_path[[t]] <- paste0(top_mut_gene$gene[mutated_gene_position])  
      }
      }
      if(!(t %in% class_spec$project_code)){ ## if cancer type is not in class then insert NA
        top_gene_in_path[[t]] <- "NA" 
      }
    }
    top_gene_in_path <- top_gene_in_path %>% do.call(rbind, .) %>% as.data.frame() %>% 'colnames<-' (n)
    clust_row <- data.frame(Class = d, row.names = n) %>% t() ## creating a row for cluster membership
    top_genes_in_path_cancer_type[[n]] <- rbind(clust_row, top_gene_in_path)
  }
  cluster_spec_enrich_table[[n]] <- top_genes_in_path_cancer_type %>% do.call(cbind, .) %>% as.data.frame() %>% 
  t() %>% as.data.frame() %>% rownames_to_column(var = "Pathway")
  }

cluster_spec_enrich_table <- cluster_spec_enrich_table %>% do.call(rbind, .) %>% as.data.frame()
save(cluster_spec_enrich_table, file = "../data/cluster_spec_enrich_table.Rda")
rm(list = ls(all.names = T))
gc()
```

```{r table-1}
load("../data/cluster_spec_enrich_table.Rda")
tumor_clusters <- fread("../data/umap_3d_coors.tsv") ## load in data

## Selecting the top 10 cancers with the largest n for table 1
n_tumor <- as.data.frame(table(tumor_clusters$project_code))
n_tumor <- n_tumor %>% arrange(desc(Freq)) %>% .[1:10, "Var1"] %>% as.character() %>% .[order(.)]

cluster_spec_enrich_table %>% .[, c("Pathway", "Class", n_tumor)] %>% kable(caption = "Table 1: Most mutated gene in class-specific pathways (enrichment score >= 0.30) for specific cancer", row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, font_size = 12) %>%
  column_spec(1, bold = T)
```
<p style="FONT-SIZE:10px; COLOR:#000000; LINE-HEIGHT:12px;">**Table 1:** The numbers in this table are completely wrong, it is just a place holder, please ignore</p>

### 2020-07-21 code/table2.Rmd
updated code to work with new cluster assignments by simon

<div style="text-align: right"> Table 2 Data </div>
```{r table-2-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_2d_3d_coors.tsv") ## load in data
load("../data/gdc_reactome_path_analysis.Rda")

tumor_clusters <- tumor_clusters %>% dplyr::select(V1, ajcc_metastasis_pathologic_pm, knn_clust)

@@

prepare_rbeta_data_p1 <- function(data, path_analysis) {
  n_tumor_in_group <- as.data.frame(table(data$ajcc_metastasis_pathologic_pm)) ## Getting the total number of tumor samples in each group
  path_analysis <- path_analysis[,names(path_analysis) %in% data$V1] ## Filter only group specific cancers
  data <- data[match(names(path_analysis), data$V1),] ## Match tumor sample ids
  table(names(path_analysis) == data$V1) ## good they match now
  ## Now we are going to build group specific cluster data
  n_tumor_in_clust <- as.data.frame(table(data$knn_clust)) ## getting total n in each cluster
  names(n_tumor_in_clust) <- c("clust", "tot_n_clust") ## rename colnames
  clust_group <- as.character(n_tumor_in_clust$clust) ## Getting the cluster numbers
  cluster_path_count <- list() ## Creating a list for the forloop
  for(n in clust_group) {
    print(n)
    df <- subset(data, knn_clust == n) ## subset cluster number
    df <- df$V1 ## getting tumor sample id
    df1 <- path_analysis[,names(path_analysis) %in% df] ## getting the tumor samples from path analysis
    df1 <- as.data.frame(rowSums(df1))
    names(df1) <- n
    cluster_path_count[[n]] <- df1
  }
  cluster_path_count <- do.call(cbind, cluster_path_count) ## Combine the list together
  rbeta_path_df <- cluster_path_count %>% rownames_to_column(var="path") 
  rbeta_path_df <- rbeta_path_df %>% gather(key = "clust", value = "mutation", -path) ## Wide to long format
  global_path_mut <- as.data.frame(rowSums(cluster_path_count))
  global_path_mut <- global_path_mut[rep(1:ncol(global_path_mut), each = max(as.numeric(as.character(rbeta_path_df$clust))))] ## Repeat column 12 times
  names(global_path_mut) <- as.character(clust_group) ## Replace names with cluster numbers
  global_path_mut <- global_path_mut %>% rownames_to_column(var="path") %>% gather( key = "clust", value = "global", -path) ## Wide to long format
  rbeta_path_df <- left_join(rbeta_path_df, global_path_mut, by = c("path", "clust")) ## left_join rbetadf and global
  rbeta_path_df <- left_join(rbeta_path_df, n_tumor_in_clust, by = "clust")
  rbeta_path_df$global_sample <- as.numeric(sum(n_tumor_in_clust$tot_n_clust)) ## Getting total tumor sample number
  names(rbeta_path_df) <- c("path", "clust", "clust_mut", "global_mut", "tot_n_clust", "global_sample")
  rbeta_path_df$tot_n_clust <- as.numeric(as.character(rbeta_path_df$tot_n_clust)) ## convert tot_n_clust numeric
  return(rbeta_path_df)
}

m0_rbeta <- prepare_rbeta_data_p1(data = m0_clust, path_analysis = gdc_reactome_path_analysis)
## Combining cluster 6/7 cluster 910 cluster 1112
combo_group <- c("6", "9", "11")
compo_group_df <- list() ## list for loop
for(n in combo_group) {
  print(n) ## first cluster in set
  m <- as.numeric(n) + 1 ## second cluster in set
  m <- as.character(m)
  df <- m0_rbeta[m0_rbeta$clust == n,] ## get rbeta info for first cluster in set
  df1 <- m0_rbeta[m0_rbeta$clust == m,] ## get rbeta info for second cluster in set
  tot_clust_mut <- df$clust_mut + df1$clust_mut ## add both clust muts
  tot_n_clust <- df$tot_n_clust + df1$tot_n_clust ## add both clust n's
  p <- paste0(n,m)
  df2 <- m0_rbeta[1:377,]
  df2$clust <- p ## replace with new clust name
  df2$clust_mut <- tot_clust_mut ## replace with new clust mut count
  df2$tot_n_clust <- tot_n_clust ## replace with new clust size
  compo_group_df[[p]] <- df2 ## store in list and continue loop
}
compo_group_df <- do.call(rbind, compo_group_df) ## rbind list
m0_rbeta <- rbind(m0_rbeta, compo_group_df) ## rbind m0_rbeta and the compo group together
save(m0_rbeta, file = "m0_rbeta.Rda")

## repeat same function above with metastasis sample
m1_rbeta <- prepare_rbeta_data_p1(data = m1_clust, path_analysis = gdc_reactome_path_analysis)
compo_group_df <- list() ## list for loop
for(n in combo_group) {
  print(n) ## first cluster in set
  m <- as.numeric(n) + 1 ## second cluster in set
  m <- as.character(m)
  df <- m0_rbeta[m0_rbeta$clust == n,] ## get rbeta info for first cluster in set
  df1 <- m0_rbeta[m0_rbeta$clust == m,] ## get rbeta info for second cluster in set
  tot_clust_mut <- df$clust_mut + df1$clust_mut ## add both clust muts
  tot_n_clust <- df$tot_n_clust + df1$tot_n_clust ## add both clust n's
  p <- paste0(n,m)
  df2 <- m0_rbeta[1:377,]
  df2$clust <- p ## replace with new clust name
  df2$clust_mut <- tot_clust_mut ## replace with new clust mut count
  df2$tot_n_clust <- tot_n_clust ## replace with new clust size
  compo_group_df[[p]] <- df2 ## store in list and continue loop
}
compo_group_df <- do.call(rbind, compo_group_df) ## rbind list
m1_rbeta <- rbind(m1_rbeta, compo_group_df) ## Combine the new groups
save(m1_rbeta, file = "m1_rbeta.Rda")

@@

save(m0_rbeta, m1_rbeta, file = "m0_m1_rbeta.Rda")
rm(list = ls(all.names = T)) ## cleaning up data
gc()
load("../data/m0_m1_rbeta.Rda")
## Metastatic (foreground) - Nonmetastatic (background) Rbeta
moderate_fg <- apply(m1_rbeta[,c("fg_mut", "fg_not_mut")], MARGIN = 1, function(x) {
  rbeta(n = 1e5, shape1=x["fg_mut"], shape2=x["fg_not_mut"])
})
moderate_bg <- apply(m0_rbeta[,c("fg_mut", "fg_not_mut")], MARGIN = 1, function(x) {
  rbeta(n = 1e5, shape1=x["fg_mut"], shape2=x["fg_not_mut"])
})

@@

m1_minus_m0_per_cluster_rbeta <- data.frame(path = m1_rbeta$path,
                                   clust = m1_rbeta$clust,
                                   probability = mean_prob,
                                   signif = sig_fg_bg[,1],
                                   difference = quant_diff[,2],
                                   lower = quant_diff[,1],
                                   upper = quant_diff[,3])
save(m1_minus_m0_per_cluster_rbeta, file = "m1_minus_m0_per_cluster_rbeta.Rda")
rm(list = ls(all.names = T))
gc()
```

>>>>

<div style="text-align: right"> Table 2 Data </div>
```{r table-2-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_3d_coors.tsv") ## load in data
load("../data/gdc_reactome_path_analysis.Rda")

tumor_clusters <- tumor_clusters %>% dplyr::select(sample_id, ajcc_metastasis_pathologic_pm, clust_knn)

@@

prepare_rbeta_data_p1 <- function(data, path_analysis) {
  n_tumor_in_group <- as.data.frame(table(data$ajcc_metastasis_pathologic_pm)) ## Getting the total number of tumor samples in each group
  path_analysis <- path_analysis[,names(path_analysis) %in% data$sample_id] ## Filter only group specific cancers
  data <- data[match(names(path_analysis), data$sample_id),] ## Match tumor sample ids
  table(names(path_analysis) == data$sample_id) ## good they match now
  ## Now we are going to build group specific cluster data
  n_tumor_in_clust <- as.data.frame(table(data$clust_knn)) ## getting total n in each cluster
  names(n_tumor_in_clust) <- c("clust", "tot_n_clust") ## rename colnames
  clust_group <- as.character(n_tumor_in_clust$clust) ## Getting the cluster numbers
  cluster_path_count <- list() ## Creating a list for the forloop
  for(n in clust_group) {
    print(n)
    df <- subset(data, clust_knn == n) ## subset cluster number
    df <- df$sample_id ## getting tumor sample id
    df1 <- path_analysis[,names(path_analysis) %in% df] ## getting the tumor samples from path analysis
    df1 <- as.data.frame(rowSums(df1))
    names(df1) <- n
    cluster_path_count[[n]] <- df1
  }
  cluster_path_count <- do.call(cbind, cluster_path_count) ## Combine the list together
  rbeta_path_df <- cluster_path_count %>% rownames_to_column(var="path") 
  rbeta_path_df <- rbeta_path_df %>% gather(key = "clust", value = "mutation", -path) ## Wide to long format
  global_path_mut <- as.data.frame(rowSums(cluster_path_count))
  global_path_mut <- global_path_mut[rep(1:ncol(global_path_mut), each = max(as.numeric(as.character(rbeta_path_df$clust))))] ## Repeat column 12 times
  names(global_path_mut) <- as.character(clust_group) ## Replace names with cluster numbers
  global_path_mut <- global_path_mut %>% rownames_to_column(var="path") %>% gather( key = "clust", value = "global", -path) ## Wide to long format
  rbeta_path_df <- left_join(rbeta_path_df, global_path_mut, by = c("path", "clust")) ## left_join rbetadf and global
  rbeta_path_df <- left_join(rbeta_path_df, n_tumor_in_clust, by = "clust")
  rbeta_path_df$global_sample <- as.numeric(sum(n_tumor_in_clust$tot_n_clust)) ## Getting total tumor sample number
  names(rbeta_path_df) <- c("path", "clust", "clust_mut", "global_mut", "tot_n_clust", "global_sample")
  rbeta_path_df$tot_n_clust <- as.numeric(as.character(rbeta_path_df$tot_n_clust)) ## convert tot_n_clust numeric
  return(rbeta_path_df)
}

m0_rbeta <- prepare_rbeta_data_p1(data = m0_clust, path_analysis = gdc_reactome_path_analysis)
save(m0_rbeta, file = "../data/m0_rbeta.Rda")

## repeat same function above with metastasis sample
m1_rbeta <- prepare_rbeta_data_p1(data = m1_clust, path_analysis = gdc_reactome_path_analysis)
save(m1_rbeta, file = "../data/m1_rbeta.Rda")

@@

save(m0_rbeta, m1_rbeta, file = "../data/m0_m1_rbeta.Rda")
rm(list = ls(all.names = T)) ## cleaning up data
gc()
load("../data/m0_m1_rbeta.Rda")
## Metastatic (foreground) - Nonmetastatic (background) Rbeta
moderate_fg <- apply(m1_rbeta[,c("fg_mut", "fg_not_mut")], MARGIN = 1, function(x) {
  rbeta(n = 1e5, shape1=x["fg_mut"], shape2=x["fg_not_mut"])
})
moderate_bg <- apply(m0_rbeta[,c("fg_mut", "fg_not_mut")], MARGIN = 1, function(x) {
  rbeta(n = 1e5, shape1=x["fg_mut"], shape2=x["fg_not_mut"])
})

@@

m1_minus_m0_per_cluster_rbeta <- data.frame(path = m1_rbeta$path,
                                   clust = m1_rbeta$clust,
                                   probability = mean_prob,
                                   signif = sig_fg_bg[,1],
                                   difference = quant_diff[,2],
                                   lower = quant_diff[,1],
                                   upper = quant_diff[,3])
save(m1_minus_m0_per_cluster_rbeta, file = "../data/m1_minus_m0_per_cluster_rbeta.Rda")
rm(list = ls(all.names = T))
gc()
```
